Robust Subspace and Higher Moments Distances for Botnet Detection

### 2019-04-27 docs/papers/2019JNCA_m-rpca/2019JNCA_m-rpca.tex
reorganization of results in files and conclusion of literature review transcription

Botnets have emerging as a key threat for cyber security, considering that Botnets are the key platform for many Internet attacks (i.e. spam, distributed denial-of-service (DDoS), identity theft, and phishing). Most of the current botnet detection approaches work only on specific botnet command and control (. Lots of efforts are directed to avoid attacks from known attackers, however attacks from one source is not real nowadays, since the attacks are usualy from distributed botnets that change their approaches frequently, in order to avoid behavioral or pattern based detecions. This changing behavior is known as adversarial model \cite{gu2008botminer, garcia2014empirical,khattak2015botflex,acarali2016survey,wang2017botnet,da2018online,Wang2018ddosbotnetssurvey}.

@@

Network anomaly detection problems are usually characterized by skewed and imbalanced data \cite{Phua2004minority,he2008learning}. Learning algorithms for imbalanced data has been a challenging research topic. The fundamental issue with the imbalanced learning problem is the ability of imbalanced data to significantly compromise the performance of most standard learning algorithms \cite{he2008learning}. He and Garcia \cite{he2008learning} present a classification of imbalanced learning methods as sampling, cost-sensitive learning, kernel-based, and active learning. Some techniques can make the data discriminative for better discriminative analytics: Data engineering (i.e. sampling \cite{he2008learning,gu2008botminer}, aggregation \cite{lakhina2005mining,gu2008botminer,callegari2011novel, garcia2014empirical, acarali2016survey}); Feature extraction (i.e. decomposition \cite{vaswani2018robust}, robust estimates \cite{zhou2017anomaly}). To face a high imbalanced data set, we adopt a sampling method based on data aggregation for discriminative feature generation, considering that data aggregation is a widely adoted approach for feature engineering \cite{garcia2014empirical, chandrashekar2014survey,acarali2016survey} and network anomaly detection \cite{lakhina2005mining, callegari2011novel, vieira2017model}.

Some approaches assume a gaussian distribute data, hoever it is not true in real world for many fields....Data skewness...

@@

Botnets are becoming one of the most serious threats to Internet security. A botnet is a network of compromised machines under the influence of malware (bot) code. The botnet is commandeered by a “botmaster” and utilized as “resource” or “platform” for attacks such as distributed denial-of-service (DDoS) attacks, and fraudulent activities such as spam, phishing, identity theft, and information exfiltration. In order for a botmaster to command a botnet, there needs to be a command and control (C&C) channel through which bots receive commands and coordinate attacks and fraudulent activities. The C&C channel is the means by which individual bots form a botnet. Centralized C&C structures using the Internet Relay Chat (IRC) protocol have been utilized by botmasters for a long time. In this architecture, each bot logs into an IRC channel, and seeks commands from the botmaster. Even today, many botnets are still designed this way. Quite a few botnets, though, have begun to use other protocols such as HTTP [8, 14, 24, 39], probably because HTTP-based C&C communications are more stealthy given that Web traffic is generally allowed in most networks \cite{gu2008botminer}.

@@

The BotHunter method was proposed by Gu \emph{et al} \cite{gu2007bothunter} to detect the infection and coordination of botnets by matching sequence model, through a correlation approach for detecting stages of the infection process.  Gu \emph{et al} \cite{gu2008botminer} presented the BotMiner, which aims to detect groups of compromised machines that are part of a botnet. BotMiner monitors communications that may suggest C&C or malicious activities, and finds a coordinated group pattern by means of clusters of similar communication activities, clusters of similar malicious activities, and performs cross cluster correlation to identify the hosts that share both similar communication patterns and similar malicious activity patterns.

@@

\cite{da2018online} (Online detection of Botnets on Network Flows using Stream Mining) Ver se podemos calcular o F1 e comparar os resutados, mas não seria justo dado que a abordagem deles nao implementa o mesmo sampling e é supervisionado

@@

\cite{Phua2004minority} (Minority Report in Fraud Detection: Classification of Skewed Data)

\cite{lakhina2005mining} (Mining anomalies using traffic feature distributions).  -Argue that the distributions of packet features (IP addresses and ports) observed in flow traces reveals both the presence and the structure of a wide range of anomalies. Using entropy as a summarization tool, we show that the analysis of feature distributions leads to significant advances on two fronts: (1) it enables highly sensitive detection of a wide range of anomalies, augmenting detections by volume-based methods, and (2) it enables automatic classification of anomalies via unsupervised learning. Anomaly diagnosis framework, comprising both an extension of the subspace method [23] to accommodate multiple data types, and an unsupervised classification technique using simple clustering algorithms. \textbf{To explain feature selection and aggregation}.

\cite{goldberg2015importance} (The importance of features for statistical anomaly detection)

@@

A. Description of CTU-13 Dataset
The CTU-13 is a dataset of botnet traffic that was captured
in the Czech Technical University [24]. It contains 13 scenarios
with various botnet types. We consider scenarios 1, 2, 6, 8, and
9 of the dataset.
• Scenario 1 corresponds to an IRC-based botnet that sent
spam for almost 6.5 hours.
• Scenario 2 (2.5-hours) is from the same botnet.
• Scenario 6 is from a botnet that scanned SMTP (Simple
Mail Transfer Protocol) servers for 2 hours and connected
to several RDP (Remote Desktop Protocol) services. Dif-
ferent with Scenarios 1 and 2, this botnet neither sent
spam nor did it attack. It’s C&C server used a proprietary
protocol.
• Scenario 8 is from a botnet that contacted a lot of Chinese
C&C hosts and received large amounts of encrypted data.
The botnet also cracked the passwords of machines during
the 19-hour attack.
• Scenario 9 corresponds to a case where 10 local hosts
were infected by a spamming botnet. More than 600 spams
were sent over 5 hours.
For all the scenarios, the authors of the CTU-13 dataset
convert the captured pcap files to NetFlows and release the
processed flows. The dataset contains ground-truth labels for
flows as follows: flows from or to the infected machines
are labeled as “botnet”; flows from or to noninfected ma-
chines are labeled as “normal”; all other flows are labeled as
“background.”

@@

Due to the importance given by F-measure to true-positive detection \cite{moustafa2019holistic}, the F-measure is the preferable measure for imbalanced data sets.

@@

	As clustering-based method, we selected K-means, which is the clustering algorithm most adopted for general purposes and for network anomaly detection \cite{gaddam2007k,moustafa2019holistic}.
	It is observed that network traffic do not belong to a Gaussian distribution \cite{benson2010network,moustafa2019holistic}, algorithms based on non-Gaussian distributions, such as GMM, has been adopted for network anomaly detection \cite{moustafa2019holistic}.

@@

Say that the data is lognormal \cite{benson2010network} and link it to skewness https://www.weibull.com/hotwire/issue47/relbasics47.htm
Test distribution fit again e present results 
Conclude this presentation.
Evaluate eigensim for CTU-13 and others algorithms for Danilo's data set.
Compare to vanilla RPCA (Eduardo is working on it).
Write a paper for JNCA (April).
Write the thesis (May and June).

>>>>

Botnets have emerging as a key threat for cyber security, considering that Botnets are the key platform for many Internet attacks (i.e. spam, distributed denial-of-service (DDoS), identity theft, and phishing). Most of the current botnet detection approaches work only on specific botnet command and control (. Lots of efforts are directed to avoid attacks from known attackers, however attacks from one source is not real nowadays, since the attacks are usualy from distributed botnets that change their approaches frequently, in order to avoid behavioral or pattern based detecions. This changing behavior is known as adversarial model \cite{gu2008botminer, garcia2014empirical,khattak2015botflex,acarali2016survey,wang2017botnet,Wang2018ddosbotnetssurvey}.

@@

Network anomaly detection problems are usually characterized by skewed and imbalanced data \cite{Phua2004minority,he2008learning,benson2010network}. Learning algorithms for imbalanced data has been a challenging research topic. The fundamental issue with the imbalanced learning problem is the ability of imbalanced data to significantly compromise the performance of most standard learning algorithms \cite{he2008learning}. He and Garcia \cite{he2008learning} present a classification of imbalanced learning methods as sampling, cost-sensitive learning, kernel-based, and active learning. Some techniques can make the data discriminative for better discriminative analytics: Data engineering (i.e. sampling \cite{he2008learning,gu2008botminer}, aggregation \cite{lakhina2005mining,gu2008botminer,callegari2011novel, garcia2014empirical, acarali2016survey}); Feature extraction (i.e. decomposition \cite{vaswani2018robust}, robust estimates \cite{zhou2017anomaly}). To face a high imbalanced data set, we adopt a sampling method based on data aggregation for discriminative feature generation, considering that data aggregation is a widely adoted approach for feature engineering \cite{garcia2014empirical, chandrashekar2014survey,acarali2016survey} and network anomaly detection \cite{lakhina2005mining, callegari2011novel, vieira2017model}. For more information about learning methods from imbalanced data, we refer to He and Garcia \cite{he2008learning}.

\cite{lakhina2005mining} (Mining anomalies using traffic feature distributions).  -Argue that the distributions of packet features (IP addresses and ports) observed in flow traces reveals both the presence and the structure of a wide range of anomalies. Using entropy as a summarization tool, we show that the analysis of feature distributions leads to significant advances on two fronts: (1) it enables highly sensitive detection of a wide range of anomalies, augmenting detections by volume-based methods, and (2) it enables automatic classification of anomalies via unsupervised learning. Anomaly diagnosis framework, comprising both an extension of the subspace method [23] to accommodate multiple data types, and an unsupervised classification technique using simple clustering algorithms. \textbf{To explain feature selection and aggregation}. Goldberg and Shan \cite{goldberg2015importance} argue that processing monitored signals into features is a more fruitful way to build anomaly detection systems than devising ever more complex statistical tests, based on experience and  on putblished real-life examples of anomaly detectors in use at eBay.

% Say that the data is lognormal \cite{benson2010network} and link it to skewness https://www.weibull.com/hotwire/issue47/relbasics47.htm
Some approaches assume a gaussian distribute data, hoever it is not true in real world for many types of network traffic, wich usually is more related to skewed and heavy-tailed distributions. Benson \emph{et al} \cite{benson2010network} conducted an empirical study of the network traffic in 10 data centers belonging to three different types of organizations, including university, enterprise, and cloud data centers. Findings of Benson \emph{et al} indicate that certain positive skewed and heavy-tailed distributions can model data center switch traffic, and highlights a difference between the data center environment and the wide area network, where the long-tailed Pareto distribution typically shows the best fit \cite{benson2010network}. This findings highligh that the skeweness and heavy-tailed distributions can impact algorithms that assum a guassian distributed data, as well as can indicate characteristics that can be exploited in order to obtain accurate classifiers for network traffic analysis.

@@

Botnets are becoming one of the most serious threats to Internet security. A botnet is a network of compromised machines under the influence of malware (bot) code. The botnet is commandeered by a “botmaster” and utilized as “resource” or “platform” for attacks such as distributed denial-of-service (DDoS) attacks, and fraudulent activities such as spam, phishing, identity theft, and information exfiltration. In order for a botmaster to command a botnet, there needs to be a command and control (C\&C) channel through which bots receive commands and coordinate attacks and fraudulent activities. The C\&C channel is the means by which individual bots form a botnet. Centralized C\&C structures using the Internet Relay Chat (IRC) protocol have been utilized by botmasters for a long time. In this architecture, each bot logs into an IRC channel, and seeks commands from the botmaster. Even today, many botnets are still designed this way. Quite a few botnets, though, have begun to use other protocols such as HTTP [8, 14, 24, 39], probably because HTTP-based C\&C communications are more stealthy given that Web traffic is generally allowed in most networks \cite{gu2008botminer}.

@@

The BotHunter method was proposed by Gu \emph{et al} \cite{gu2007bothunter} to detect the infection and coordination of botnets by matching sequence model, through a correlation approach for detecting stages of the infection process.  Gu \emph{et al} \cite{gu2008botminer} presented the BotMiner, which aims to detect groups of compromised machines that are part of a botnet. BotMiner monitors communications that may suggest C\&C or malicious activities, and finds a coordinated group pattern by means of clusters of similar communication activities, clusters of similar malicious activities, and performs cross cluster correlation to identify the hosts that share both similar communication patterns and similar malicious activity patterns.

@@

% \cite{da2018online} proposes an approach using the Very Fast Decision Tree, a classification algorithm used on stream mining that can learn incrementally when needed, to identify botnets by observing network flows. When evaluating the approach on multiple scenarios with different botnets, we were able to achieve high performance metrics on the majority of scenarios, while using a significantly low number of labelled instances. It discard scenarios 5 and 6. They does not compute F1, but it is easy to compute it from precision and recall provided.

@@

\cite{zhou2017anomaly} (Anomaly detection with robust deep autoencoders)

@@

A. Description of CTU-13 Dataset
The CTU-13 is a dataset of botnet traffic that was captured
in the Czech Technical University [24]. It contains 13 scenarios
with various botnet types. We consider scenarios 1, 2, 6, 8, and
9 of the dataset.
• Scenario 1 corresponds to an IRC-based botnet that sent
spam for almost 6.5 hours.
• Scenario 2 (2.5-hours) is from the same botnet.
• Scenario 6 is from a botnet that scanned SMTP (Simple
Mail Transfer Protocol) servers for 2 hours and connected
to several RDP (Remote Desktop Protocol) services. Dif-
ferent with Scenarios 1 and 2, this botnet neither sent
spam nor did it attack. It’s C\&C server used a proprietary
protocol.
• Scenario 8 is from a botnet that contacted a lot of Chinese
C\&C hosts and received large amounts of encrypted data.
The botnet also cracked the passwords of machines during
the 19-hour attack.
• Scenario 9 corresponds to a case where 10 local hosts
were infected by a spamming botnet. More than 600 spams
were sent over 5 hours.
For all the scenarios, the authors of the CTU-13 dataset
convert the captured pcap files to NetFlows and release the
processed flows. The dataset contains ground-truth labels for
flows as follows: flows from or to the infected machines
are labeled as “botnet”; flows from or to noninfected ma-
chines are labeled as “normal”; all other flows are labeled as
“background.”

@@

Due to the importance given by F-measure to true-positive detection \cite{powers2011evaluation,moustafa2019holistic}, the F-measure is the preferable measure for imbalanced data sets.

@@

Why GMM?
As clustering-based method, we selected K-means, which is the clustering algorithm most adopted for general purposes and for network anomaly detection \cite{gaddam2007k,moustafa2019holistic}.
It is observed that network traffic do not belong to a Gaussian distribution \cite{benson2010network,moustafa2019holistic}, algorithms based on non-Gaussian distributions, such as GMM, has been adopted for network anomaly detection \cite{moustafa2019holistic}.

@@

Test distribution fit again e present results 
Conclude this presentation.
Evaluate eigensim for CTU-13 and others algorithms for Danilo's data set.
Compare to vanilla RPCA (Eduardo is working on it).
Write a paper for JNCA (April).
Write the thesis (May and June).
